---
title: "Tesla Stock Volatility Analysis using GARCH Models"
output: html_document
date: "2026-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Introduction

This analysis examines the volatility dynamics of Tesla stock returns using GARCH models.
The objective is to:

* Compute and analyze daily log returns
* Test for volatility clustering (heteroskedasticity)
* Fit a GARCH(1,1) model
* Evaluate model diagnostics
* Forecast future volatility

---

# 1. Load Libraries

```{r}
library(tidyverse)
library(lubridate)
library(FinTS)
library(rugarch)
library(xts)
```

---

# 2. Data Import and Preparation

```{r}
tesla <- read.csv("C:/Users/admin/Desktop/New folder/su work/1.2/timeseries/Garch_Assignment/TSLA1.csv")

head(tesla)
tail(tesla)
```

### Convert Date and Compute Log Returns

```{r}
tesla$Date <- as.Date(tesla$Date)

tesla <- tesla %>%
  arrange(Date) %>%
  mutate(
    log_return = 100 * (log(Adj.Close) - log(lag(Adj.Close)))
  ) %>%
  drop_na()

summary(tesla$log_return)
```

---

# 3. Exploratory Analysis

### Plot Daily Log Returns

```{r}
plot(tesla$Date, tesla$log_return, type = "l",
     main = "Tesla Daily Log Returns",
     xlab = "Date",
     ylab = "Log Returns (%)")
```

### Squared Returns (Volatility Clustering)

```{r}
plot(tesla$log_return^2, type = "l",
     main = "Squared Returns (Volatility Clustering)",
     ylab = "Squared Returns")
```

---

# 4. Test for ARCH Effects

```{r}
ArchTest(tesla$log_return, lags = 12)
```

**Interpretation:**
A significant p-value indicates the presence of ARCH effects (time-varying volatility), justifying the use of GARCH models.

---

# 5. GARCH Model Specification and Estimation

```{r}
spec_garch <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
  mean.model     = list(armaOrder = c(0,0), include.mean = TRUE),
  distribution.model = "std"   # Student-t distribution
)

fit_garch <- ugarchfit(
  spec = spec_garch,
  data = tesla$log_return
)

show(fit_garch)
```

---

# 6. Conditional Volatility Visualization

```{r}
tesla_xts <- xts(tesla$log_return, order.by = tesla$Date)
volatility <- sigma(fit_garch)

plot(tesla$Date, volatility, type = "l",
     main = "Estimated Conditional Volatility (GARCH 1,1)",
     xlab = "Date",
     ylab = "Volatility (%)")
```

---

# 7. Model Diagnostics

### Standardized Residuals

```{r}
residuals <- residuals(fit_garch, standardize = TRUE)
```

### Ljung-Box Test

```{r}
Box.test(residuals^2, lag = 20, type = "Ljung-Box")
```

### ARCH Test on Residuals

```{r}
ArchTest(residuals, lags = 12)
```

### Residual Analysis Plots

```{r}
par(mfrow = c(2,2))

acf(residuals, main = "ACF of Residuals")
acf(residuals^2, main = "ACF of Squared Residuals")

hist(residuals, breaks = 50,
     main = "Histogram of Residuals",
     xlab = "Residuals")

qqnorm(residuals)
qqline(residuals, col = "red")

par(mfrow = c(1,1))
```

---

# 8. Volatility Forecasting

```{r}
forecast <- ugarchforecast(fit_garch, n.ahead = 10)

vol_forecast <- sigma(forecast)

last_date <- max(tesla$Date)
forecast_dates <- seq(last_date + 1, by = "days", length.out = 10)

vol_forecast_xts <- xts(vol_forecast, order.by = forecast_dates)

vol_forecast_xts
```

---

# Conclusion

The analysis confirms the presence of volatility clustering in Tesla stock returns.
The GARCH(1,1) model effectively captures time-varying volatility and provides reasonable short-term forecasts.

Such models are essential in financial risk management, portfolio optimization, and derivative pricing.

```
